Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-08 introduction.tex
Ajout exemples

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete division of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear repreentation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events. Throughout the paper we will
illustrate our framework with examples taken for this motivating application.

@@

% First one general edit-distance is defined by a weighted word
% transducer~\cite{Mohri}
% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
% %dealing with infinite set of input symbols...
% value in a semiring...
\florent{expressiveness: VPA have restricted equality test.
        comparable to pebble automata? $\to$ conclusion}
The main contributions of the paper are:
($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
generalizing the corresponding classes of symbolic and weighted models,
%for the weighted computation on (nested) words over infinite alphabets;
($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
% a framework for parsing over infinite alphabets,
and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets,
the keys to which are
($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
and ($iii.b$)~the use, convenient in this context,
of nested words, and \SWVPA,
instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
%
%Finally, Section~\ref{sec:transcription} describes
%the implemented application %of this approach
%to automated music transcription
%that motivated this study.

%based these models and on generic def. of distances;
%the use of VPA for parsing, word instead of trees, use comparison input-output.
%and best-search

%In Section~\ref{section:SWA} we introduce \SWA and \SWT.
%Then \SWVPA are defined in Section~\ref{sec:SWVPA},
%where a polynomial 1-best algorithm is described that can be use to solve
%Symbolic Weighted Parsing, as described Section~\ref{sec:parsing}.
%Finally, in Section~\ref{sec:transcription}, we present an application
%of this approach to automated music transcription that has been implemented.

>>>>

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete division of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear representation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events.

@@

% First one general edit-distance is defined by a weighted word
% transducer~\cite{Mohri}
% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
% %dealing with infinite set of input symbols...
% value in a semiring...
\florent{expressiveness: VPA have restricted equality test.
        comparable to pebble automata? $\to$ conclusion}
The main contributions of the paper are:
($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
generalizing the corresponding classes of symbolic and weighted models,
%for the weighted computation on (nested) words over infinite alphabets;
($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
% a framework for parsing over infinite alphabets,
and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets,
the keys to which are
($iii.a$)~the \SWT-based definition of generic edit distances between input and output (yield) words,
and ($iii.b$)~the use, convenient in this context,
of nested words, and \SWVPA,
instead of syntax trees and grammars. %(\S~\ref{sec:trees}).

\begin{example}[Running example]
Throughout the paper we
illustrate our framework with music transcription examples: Given
a \emph{timeline} of musical events with arbitrary timestamps as input, parse it
into a structured music score. In our example, input events
are pairs $\< \eta, \tau>$ made of
a symbol $\eta \in \Sigma$, where $\Sigma$ stands for the
set of MIDI message symbols~\cite{?}
and  $\tau \in \mathbb{Q}$ is a timestamp. The output of parsing
is a representation of the sequence in
Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where event symbols belong to the domain $\Delta$
of \emph{pitches} (e.g., A4, G5, etc.), temporal
information is encoded as \emph{durations} (whole \musWhole$,
$quarter, $\musQuarter$, eight $\musEighth$, etc), and notes are grouped in
high-level structures (beams, measures, tuplets). The following inputs
will be used:
\begin{enumerate}
  \item $I_1  = [ <e_1, 0.07>,<e_2, 0.72>,<e_3, 0.91>]$, over interval $[0,1[$
  \item $I_2  = [ <e_3, 1.05>,<e_4, 1.36>,<e_5, 1.71>]$, over interval $[1,2[$
\end{enumerate}
There exists many possible parsings of $I_1 \cup I_2$ in music notation, among
which
\includegraphics[scale=0.20]{pictures/score5.png}
and \includegraphics[scale=0.20]{pictures/score4.png}.  Wheighted
parsing associates a cost to each solution, and our framework
aims at selecting the best one with respect to this cost.
\endex
\end{example}
